---
title: "Swarm Art in the Browser with boids4R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Swarm Art in the Browser with boids4R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
if (file.exists("../DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else {
  library(ggWebGL)
}

boids4r_available <- requireNamespace("boids4R", quietly = TRUE)
```

# Swarm Art in the Browser

This article shows how `ggWebGL` can render swarm art in the browser from
renderer-neutral `boids4R` simulations. The visual idea is generative rather
than statistical: compact schools, predator avoidance, obstacle corridors,
murmurations, and mixed-species flocks become animated point/vector timelines
that can be panned, zoomed, scrubbed, and inspected in WebGL.

The examples are deliberately browser-native. `boids4R` computes the flocking
frames, and `ggWebGL` turns those frames into animated primitives with shader
and camera controls. The same handoff can be used for gallery pieces, teaching
demonstrations, or stress tests of timeline rendering.

# Boundary

The ownership split is deliberately narrow:

- `boids4R` owns simulation semantics: flocking state, rules, scenarios, and
  frame export
- `boids4R::as_ggwebgl_spec()` translates simulation frames into renderer
  primitives
- `ggWebGL` owns WebGL rendering: the htmlwidget, point and vector primitives,
  shader choice, timeline controls, hover, pan, zoom, and 3D orbit interaction

`boids4R` is a suggested package for `ggWebGL`. When it is unavailable, this
vignette still builds and reports that the live animation widgets were skipped.

```{r}
if (!boids4r_available) {
  cat("boids4R is unavailable, so live boids animations are skipped in this vignette.\n")
} else {
  cat("boids4R is available; live boids animations will be rendered below.\n")
}
```

# Scenario Gallery: Swarm Art Motifs

These examples mirror the `boids4R` scenario gallery: compact schools, obstacle
corridors, predator avoidance, murmurations, and mixed-species 3D flocks. Each
widget is built from a renderer-neutral `boids_simulation` object and animated
by the `ggWebGL` timeline.

```{r}
if (!boids4r_available) {
  cat("Scenario gallery widgets skipped.\n")
} else {
  scenario_gallery <- list(
    schooling_2d = boids4R::boids_scenario(
      "schooling_2d",
      n = 180L,
      steps = 80L,
      record_every = 2L,
      seed = 111L
    ),
    obstacle_corridor_2d = boids4R::boids_scenario(
      "obstacle_corridor_2d",
      n = 160L,
      steps = 90L,
      record_every = 2L,
      seed = 112L
    ),
    predator_avoidance_2d = boids4R::boids_scenario(
      "predator_avoidance_2d",
      n = 170L,
      steps = 90L,
      record_every = 2L,
      seed = 113L
    ),
    murmuration_3d = boids4R::boids_scenario(
      "murmuration_3d",
      n = 240L,
      steps = 90L,
      record_every = 3L,
      seed = 114L
    ),
    mixed_species_3d = boids4R::boids_scenario(
      "mixed_species_3d",
      n = 210L,
      steps = 90L,
      record_every = 3L,
      seed = 115L
    )
  )

  scenario_widgets <- lapply(names(scenario_gallery), function(name) {
    sim <- scenario_gallery[[name]]
    spec <- boids4R::as_ggwebgl_spec(
      sim,
      vector_every = if (identical(sim$dimension, "3d")) 16L else 12L,
      vector_scale = if (identical(sim$dimension, "3d")) 0.11 else 0.13,
      shader = "density_splat"
    )
    spec$labels$title <- paste("boids4R", name)
    spec$render$timeline$autoplay <- TRUE
    spec$render$timeline$loop <- TRUE
    spec$render$timeline$speed <- 1.4
    ggWebGL::ggWebGL(spec, height = if (identical(sim$dimension, "3d")) 540 else 500)
  })
  names(scenario_widgets) <- names(scenario_gallery)
}
```

## Schooling 2D

```{r}
if (!boids4r_available) {
  cat("Schooling widget skipped.\n")
} else {
  scenario_widgets$schooling_2d
}
```

## Obstacle Corridor 2D

```{r}
if (!boids4r_available) {
  cat("Obstacle corridor widget skipped.\n")
} else {
  scenario_widgets$obstacle_corridor_2d
}
```

## Predator Avoidance 2D

```{r}
if (!boids4r_available) {
  cat("Predator avoidance widget skipped.\n")
} else {
  scenario_widgets$predator_avoidance_2d
}
```

## Murmuration 3D

```{r}
if (!boids4r_available) {
  cat("Murmuration widget skipped.\n")
} else {
  scenario_widgets$murmuration_3d
}
```

## Mixed Species 3D

```{r}
if (!boids4r_available) {
  cat("Mixed-species widget skipped.\n")
} else {
  scenario_widgets$mixed_species_3d
}
```

# Custom Workflow Animations

The custom workflow builds a corridor from low-level `boids4R` constructors,
then compares a baseline run with a stronger obstacle/predator avoidance run.

```{r}
if (!boids4r_available) {
  cat("Custom workflow widgets skipped.\n")
} else {
  bounds <- matrix(
    c(-2.4, -1.35, 2.4, 1.35),
    ncol = 2,
    dimnames = list(c("x", "y"), c("min", "max"))
  )

  n_school <- 96L
  n_scout <- 32L
  n_boids <- n_school + n_scout
  school_axis <- seq(0, 1, length.out = n_school)
  scout_axis <- seq(0, 1, length.out = n_scout)
  positions <- rbind(
    cbind(-2.18 + 0.83 * school_axis, -0.70 + 0.95 * abs(sin(pi * school_axis))),
    cbind(-2.22 + 0.77 * scout_axis, 0.28 + 0.64 * abs(cos(pi * scout_axis)))
  )
  velocity_phase <- seq(0, 2 * pi, length.out = n_boids)
  velocities <- cbind(
    0.35 + 0.20 * cos(velocity_phase),
    0.08 * sin(velocity_phase)
  )

  custom_state <- boids4R::boids_state(
    n_boids,
    "2d",
    bounds = bounds,
    positions = positions,
    velocities = velocities,
    species = c(rep("school", n_school), rep("scout", n_scout))
  )

  custom_world <- boids4R::boids_world(
    "2d",
    bounds = bounds,
    boundary = "reflect",
    obstacles = data.frame(
      x = c(-0.82, -0.05, 0.72),
      y = c(0.42, -0.36, 0.48),
      radius = c(0.30, 0.36, 0.31)
    ),
    predators = data.frame(
      x = -0.25,
      y = 0.92,
      radius = 0.58,
      strength = 1.2
    ),
    attractors = data.frame(
      x = 2.08,
      y = -0.86,
      strength = 0.95
    )
  )

  baseline_params <- boids4R::boids_params(
    "2d",
    separation_weight = 1.35,
    alignment_weight = 0.94,
    cohesion_weight = 0.62,
    obstacle_weight = 2.5,
    predator_weight = 2.3,
    goal_weight = 0.16,
    max_speed = 1.18,
    max_force = 0.12,
    noise = 0.001
  )

  avoidance_params <- boids4R::boids_params(
    "2d",
    separation_weight = 1.35,
    alignment_weight = 0.94,
    cohesion_weight = 0.62,
    obstacle_weight = 2.8,
    predator_weight = 3.2,
    goal_weight = 0.20,
    max_speed = 1.18,
    max_force = 0.12,
    noise = 0.001
  )

  custom_runs <- list(
    baseline = boids4R::simulate_boids(
      custom_state,
      custom_world,
      baseline_params,
      steps = 95L,
      record_every = 2L,
      seed = 221L
    ),
    stronger_avoidance = boids4R::simulate_boids(
      custom_state,
      custom_world,
      avoidance_params,
      steps = 95L,
      record_every = 2L,
      seed = 222L
    )
  )

  custom_widgets <- lapply(names(custom_runs), function(name) {
    spec <- boids4R::as_ggwebgl_spec(
      custom_runs[[name]],
      vector_every = 10L,
      vector_scale = 0.14,
      shader = "density_splat"
    )
    spec$labels$title <- paste("boids4R custom corridor:", name)
    spec$render$timeline$autoplay <- TRUE
    spec$render$timeline$loop <- TRUE
    spec$render$timeline$speed <- 1.5
    ggWebGL::ggWebGL(spec, height = 500)
  })
  names(custom_widgets) <- names(custom_runs)
}
```

## Baseline Corridor

```{r}
if (!boids4r_available) {
  cat("Baseline corridor widget skipped.\n")
} else {
  custom_widgets$baseline
}
```

## Stronger Avoidance Corridor

```{r}
if (!boids4r_available) {
  cat("Stronger avoidance corridor widget skipped.\n")
} else {
  custom_widgets$stronger_avoidance
}
```

# Regeneration

The same pattern can be used from an installed package:

```{r eval = FALSE}
sim <- boids4R::boids_scenario("mixed_species_3d", n = 210, steps = 90, seed = 115)
spec <- boids4R::as_ggwebgl_spec(sim, vector_every = 16, shader = "density_splat")
spec$render$timeline$autoplay <- TRUE
ggWebGL::ggWebGL(spec, height = 540)
```
