---
title: "Experimental 3D, mesh, and surface WebGL layers"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Experimental 3D, mesh, and surface WebGL layers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
ggwebgl_truthy <- function(x) {
  tolower(x) %in% c("1", "true", "yes", "y")
}

ggwebgl_ci_vars <- c(
  "CI",
  "GITHUB_ACTIONS",
  "GITLAB_CI",
  "BUILDKITE",
  "TRAVIS",
  "APPVEYOR",
  "CIRCLECI",
  "JENKINS_URL"
)
ggwebgl_is_ci <- any(vapply(Sys.getenv(ggwebgl_ci_vars), ggwebgl_truthy, logical(1)))
ggwebgl_is_check <- nzchar(Sys.getenv("_R_CHECK_PACKAGE_NAME_"))
ggwebgl_eval_examples <- !ggwebgl_is_ci &&
  !ggwebgl_is_check &&
  (
    ggwebgl_truthy(Sys.getenv("NOT_CRAN")) ||
      ggwebgl_truthy(Sys.getenv("GGWEBGL_EVAL_COVERAGE_VIGNETTE"))
  )

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = ggwebgl_eval_examples
)

library(ggplot2)
library(ggWebGL)
```

The 3D and indexed-geometry APIs are exported for browser-side WebGL rendering,
but they remain experimental because camera, picking, and material behavior may
still evolve. Evaluation is disabled during CRAN, package checks, and CI unless
explicitly enabled with `GGWEBGL_EVAL_COVERAGE_VIGNETTE=true` or
`NOT_CRAN=true`.

```{r coverage-eval-note, echo = FALSE, results = "asis", eval = TRUE}
if (!ggwebgl_eval_examples) {
  cat(
    "> Example chunks are shown but not evaluated in this build. ",
    "Set `GGWEBGL_EVAL_COVERAGE_VIGNETTE=true` or `NOT_CRAN=true` ",
    "to evaluate them during a local non-CI render.\n",
    sep = ""
  )
}
```

# Experimental 3D, Mesh, and Surface Layers

Use `coord_webgl_3d()` to opt into a 3D camera path for grammar-style plots.
The same scene contract can also be produced through renderer-ready mesh and
surface helpers.

## Applet: 3D helix path

```{r three-d-helix-path}
helix_t <- seq(0, 4 * pi, length.out = 80)
helix <- data.frame(
  x = cos(helix_t),
  y = sin(helix_t),
  z = helix_t / max(helix_t),
  time = helix_t,
  group = "helix"
)

p <- ggplot(helix, aes(x, y, z = z, group = group, time = time)) +
  geom_path3d_webgl(colour = "#2563eb") +
  coord_webgl_3d() +
  labs(title = "3D helix path")

ggplot_webgl(p, height = 420)
```

## Applet: Structured surface

Structured surfaces use a regular grid.

```{r structured-surface}
surface_grid <- expand.grid(
  x = seq(-1, 1, length.out = 12),
  y = seq(-1, 1, length.out = 12),
  KEEP.OUT.ATTRS = FALSE
)
surface_grid$z <- with(surface_grid, exp(-(x^2 + y^2)))

p <- ggplot(surface_grid, aes(x, y, z = z, fill = z)) +
  geom_surface_webgl(shading = "surface_height_colormap") +
  coord_webgl_3d() +
  labs(title = "Structured surface")

ggplot_webgl(p, height = 420)
```

## Applet: Unstructured mesh

Unstructured meshes can be supplied with explicit vertices and triangle indices.

```{r unstructured-mesh}
mesh_vertices <- data.frame(
  x = c(0, 1, 0, 0),
  y = c(0, 0, 1, 0),
  z = c(0, 0, 0, 1),
  scalar = c(0, 1, 1, 0.5),
  i = c(1, 1, 1, 2),
  j = c(2, 2, 3, 3),
  k = c(3, 4, 4, 4)
)

p <- ggplot(mesh_vertices, aes(x, y, z = z, i = i, j = j, k = k, scalar = scalar)) +
  geom_mesh_webgl(shading = "mesh_scalar_colormap") +
  coord_webgl_3d() +
  labs(title = "Unstructured mesh")

ggplot_webgl(p, height = 420)
```

# 3D View and Camera Notes

- `coord_webgl_3d()` requests a 3D coordinate system and an orbit-style camera.
- Surface layers are structured-grid primitives; arbitrary triangle data belongs
  in mesh layers.
- The 3D APIs are experimental and should be treated as browser-side rendering
  paths, not as native GPU preprocessing pipelines.
