---
title: "Interactive benchmark metrics with ggWebGL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive benchmark metrics with ggWebGL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

if (file.exists("DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(".", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else if (file.exists("../DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else {
  library(ggWebGL)
}

benchmark_candidates <- c(
  "inst/benchmarks/benchmark-scene-types.R",
  file.path("..", "inst", "benchmarks", "benchmark-scene-types.R"),
  system.file("benchmarks", "benchmark-scene-types.R", package = "ggWebGL")
)
benchmark_candidates <- benchmark_candidates[nzchar(benchmark_candidates) & file.exists(benchmark_candidates)]
if (!length(benchmark_candidates)) {
  stop("Could not find benchmark-scene-types.R")
}
sys.source(benchmark_candidates[[1L]], envir = knitr::knit_global())

workflow_candidates <- c(
  "inst/examples/htmlwidget/workflow-comparison.R",
  file.path("..", "inst", "examples", "htmlwidget", "workflow-comparison.R"),
  system.file("examples", "htmlwidget", "workflow-comparison.R", package = "ggWebGL")
)
workflow_candidates <- workflow_candidates[nzchar(workflow_candidates) & file.exists(workflow_candidates)]
if (!length(workflow_candidates)) {
  stop("Could not find workflow-comparison.R")
}
sys.source(workflow_candidates[[1L]], envir = knitr::knit_global())
```

# Purpose

`ggWebGL` includes manual benchmark helpers for measuring browser rendering
behavior across representative scene types. These helpers are diagnostic: rows
without browser metadata are not complete browser measurements, and fixed
performance numbers should not be reported from incomplete rows.

The benchmark scripts are not run automatically by package checks. They write
only to an explicit `output_dir` or `tempdir()`.

# Metric Schema

The scene benchmark helper records a stable set of columns so local browser
runs can be compared without changing downstream analysis code.

```{r benchmark-columns}
benchmark_scene_columns()
```

The most important groups are:

- scene identity and package metadata;
- payload and artifact sizes;
- startup and browser frame timings;
- browser, device, GPU, and pixel-size metadata;
- artifact paths and completion status.

# Manual Dense Embedding Run

The dense embedding helper builds a deterministic million-point scene using
compact point transport, progressive upload metadata, the `density_splat`
shader, and rectangular brush selection.

```{r dense-benchmark-command, eval = FALSE}
source(system.file("examples", "htmlwidget", "million-point-embedding.R", package = "ggWebGL"))
run_manual_million_point_embedding()
```

# Scene-Type Metrics

Use `benchmark_scene_types()` to produce a CSV with one row per scene. Browser
metrics are populated only when a browser capture path is explicitly requested
and available.

```{r scene-type-command, eval = FALSE}
source(system.file("benchmarks", "benchmark-scene-types.R", package = "ggWebGL"))
metrics <- benchmark_scene_types(
  output_dir = tempdir(),
  scenes = c("embedding", "trajectories", "surface_mesh", "workflow"),
  point_count = 1000000L,
  include_browser = FALSE
)
metrics[, c("scene_id", "serialized_bytes", "artifact_bytes", "status")]
```

# Workflow Comparison

A useful smoke check is whether the same data mapping can be inspected both as a
static `ggplot2` graphic and as a browser-native WebGL widget. The reduced
example below uses deterministic in-memory data.

```{r workflow-static, fig.width=6, fig.height=4}
workflow_comparison_plot(800L)
```

```{r workflow-webgl, out.width='100%'}
workflow_comparison_widget(800L, height = 360)
```

# Interpreting Results

Rows with `status = "browser_skipped"` are useful for checking payload size and
widget artifact size only. Browser timing fields should be interpreted only
when the row includes browser, GPU, device, and pixel-size metadata from the
machine that performed the run.
