---
title: "Real-Data Evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real-Data Evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(ggWebGL)
helper_path <- system.file("examples", "real", "real-data-helpers.R", package = "ggWebGL")
source(helper_path, local = knitr::knit_global())
real_examples <- real_data_plots()
real_info <- real_data_metadata()
```

# Real-Data Lens

This vignette complements the synthetic showcase with real offline datasets that
exercise the renderer on:

- a raster scalar field
- observed storm trajectories
- a dense 10,000-point projection
- fixed-scale faceted dense point panels

The goal is evidence rather than maximal detail. Every example is built from
packaged data that is available without network access.

For network-enabled environments, `inst/scripts/download_mnist_embedding.R`
provides an optional path to a larger MNIST-style embedding export. The
packaged vignette keeps the offline dense embedding so the article builds
reliably everywhere.

# Example 1: Topographic Raster Field

**Why this example matters.** This is the raster path on a real elevation grid.
It exists to demonstrate that `geom_raster_webgl()` is now a rendered raster
field, not only an exported placeholder.

**What to inspect.** Terrain structure should remain continuous during pan and
zoom, and interpolation should avoid visible block artifacts in the default
example.

```{r volcano-dem, out.width='100%'}
real_examples$volcano_dem
```

```{r volcano-demgl, out.width='100%'}
ggplot_webgl(real_examples$volcano_dem+theme_webgl(shader = "default"), height = 620)
```

```{r volcano-demgl2, out.width='100%'}
ggplot_webgl(real_examples$volcano_dem, height = 620)
```

# Example 2: Observed Storm Trajectories

**Why this example matters.** This is the real trajectory case. It demonstrates
the package on geographic path bundles with real timestamps, wind speeds, and
pressure measurements behind the rendered lines.

**What to inspect.** Age shading and endpoint emphasis should make direction and
final position legible across multiple storms.

```{r storm-tracks, out.width='100%'}
real_examples$storm_tracks
```

```{r storm-tracksgl, out.width='100%'}
ggplot_webgl(real_examples$storm_tracks+theme_webgl(shader = "default"), height = 620)
```

```{r storm-tracksgl2, out.width='100%'}
ggplot_webgl(real_examples$storm_tracks, height = 620)
```

# Example 3: Dense Real-World Embedding

**Why this example matters.** This is the dense point-cloud case using a real
10,000-point feature projection. It tests whether the point renderer remains
legible on a real dense distribution instead of only on synthetic blobs.

**What to inspect.** Look for high-density accumulation without collapse into an
undifferentiated cloud.

```{r dense-embedding, out.width='100%'}
real_examples$dense_embedding
```

```{r dense-embeddinggl, out.width='100%'}
ggplot_webgl(real_examples$dense_embedding+theme_webgl(shader = "default"), height = 620)
```

```{r dense-embeddinggl2, out.width='100%'}
ggplot_webgl(real_examples$dense_embedding, height = 620)
```

# Example 4: Fixed-Scale Faceted Panels

**Why this example matters.** This is the facet-support case. It exercises the
new multi-panel layout on dense real point data.

**What to inspect.** Each panel should clip correctly, preserve its own
interaction state, and remain comparable because scales stay fixed across
panels.

```{r faceted-embedding, out.width='100%'}
real_examples$faceted_embedding
```

```{r faceted-embeddinggl, out.width='100%'}
ggplot_webgl(real_examples$faceted_embedding+theme_webgl(shader = "default"), height = 720)
```

```{r faceted-embeddinggl2, out.width='100%'}
ggplot_webgl(real_examples$faceted_embedding, height = 720)
```

# Real-Data Demo

The same real-data scenes are available in a Shiny app:

```r
source(system.file("examples", "shiny", "real-data-demo.R", package = "ggWebGL"))
```

And as standalone HTML exports:

```r
source(system.file("examples", "htmlwidget", "real-data-gallery.R", package = "ggWebGL"))
export_real_data_gallery()
```
