---
title: "Renderer Showcase Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Renderer Showcase Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(ggWebGL)
helper_path <- system.file("examples", "showcase", "showcase-helpers.R", package = "ggWebGL")
source(helper_path, local = knitr::knit_global())
showcase_info <- showcase_metadata()
showcase_examples <- showcase_plots(detail = "standard")
```

# Showcase Lens

This vignette assembles several scenarios meant to answer the kind of question a
package user is likely to ask:

- Does the package support visually compelling, research-relevant examples?
- Can it express both generative-machine-learning and scientific-visualization
  workflows through a grammar familiar to R users?
- Is the current rendering surface broad enough to support a larger WebGL
  workflow?

The current package renders dense 2D points and lines in WebGL. The examples
below stay within that supported subset on purpose.

The vignette uses the `standard` preset to keep article build time and page
weight reasonable. The standalone gallery exporter uses
`detail = "high_detail"` by default.

The examples also use the current shader surface deliberately:

- `density_splat` for dense point fields
- `trajectory_age` for line and path bundles
- `hover` inspection in addition to pan and zoom

# Example 1: Latent-Space Population Structure

**Why this example matters.** This is the dense embedding case. It demonstrates
whether ggWebGL can display cluster cores and transitional bridges in a way that
feels closer to an interactive graphics system than to a static statistical
scatterplot.

**What to inspect.** The transitional bridge should read as a continuous
structure rather than as noise between clusters.

**Rendering mode.** This example uses `shader = "density_splat"` to make dense
regions accumulate rather than read as a uniform cloud of isolated marks.

```{r latent-cloud, out.width='100%'}
showcase_examples$latent_cloud
```

```{r latent-cloudgl, out.width='100%'}
ggplot_webgl(showcase_examples$latent_cloud+theme_webgl(shader = "default", height = 630))
```

```{r latent-cloudgl2, out.width='100%'}
ggplot_webgl(showcase_examples$latent_cloud, height = 630)
```

# Example 2: Diffusion-Style Denoising Trajectories

**Why this example matters.** This is the generative-ML example. It shows many
simultaneous trajectories collapsing toward a small set of modes, which is the
kind of dynamic structure a package user might expect from an interactive
renderer demonstration.

**What to inspect.** Look for smooth bundles, endpoint concentration, and clear
distinction between target modes.

**Rendering mode.** This example uses `shader = "trajectory_age"` so later
steps are visually emphasized along each denoising path.

```{r diffusion-paths, out.width='100%'}
showcase_examples$diffusion_paths
```

```{r diffusion-pathsgl, out.width='100%'}
ggplot_webgl(showcase_examples$diffusion_paths+theme_webgl(shader = "default", height = 630), height = 630)
```

```{r diffusion-pathsgl2, out.width='100%'}
ggplot_webgl(showcase_examples$diffusion_paths, height = 630)
```

# Example 3: Nonlinear Phase Portrait

**Why this example matters.** This positions ggWebGL as a scientific
visualization tool rather than only as an exploratory embedding viewer. The
value proposition here is that many dynamical trajectories can remain legible in
an interactive browser-native plot.

**What to inspect.** The repeated loops and eventual convergence patterns should
be visible without resorting to raster precomputation.

**Rendering mode.** This example also uses `shader = "trajectory_age"` so the
flow direction remains legible in dense orbit bundles.

```{r phase-portrait, out.width='100%'}
showcase_examples$phase_portrait
```

```{r phase-portraitgl, out.width='100%'}
ggplot_webgl(showcase_examples$phase_portrait+theme_webgl(shader = "default"), height = 630)
```

```{r phase-portraitgl2, out.width='100%'}
ggplot_webgl(showcase_examples$phase_portrait, height = 630)
```

# Example 4: Optimization Paths on a Surrogate Loss Landscape

**Why this example matters.** This is a mixed-layer example. Dense points
provide a surrogate landscape while line layers show optimizer paths. It is
useful because it demonstrates composition, not only raw primitive throughput.

**What to inspect.** The background field provides context; the traces show
method-specific descent behaviour and final convergence points.

**Rendering mode.** This example uses `shader = "density_splat"` for the dense
point field while leaving the optimizer traces crisp and readable.

```{r loss-landscape, out.width='100%'}
showcase_examples$loss_landscape
```

```{r loss-landscapegl, out.width='100%'}
ggplot_webgl(showcase_examples$loss_landscape+theme_webgl(shader = "default"), height = 630)
```

```{r loss-landscapegl2, out.width='100%'}
ggplot_webgl(showcase_examples$loss_landscape, height = 630)
```

```{r loss-landscapegl3, out.width='100%'}
ggplot_webgl(showcase_examples$loss_landscape+theme_webgl(shader = "trajectory_age"), height = 630)
```

```{r loss-landscapegl4, out.width='100%'}
ggplot_webgl(showcase_examples$loss_landscape+theme_webgl(shader = "trajectory_age_glow"), height = 630)
```


# Shiny Demo

The same scenarios are available in an interactive showcase Shiny app:

```r
source(system.file("examples", "shiny", "showcase-demo.R", package = "ggWebGL"))
```

The standalone gallery export uses the denser preset:

```r
source(system.file("examples", "htmlwidget", "renderer-showcase-gallery.R", package = "ggWebGL"))
export_renderer_showcase_gallery(detail = "high_detail")
```

This app pairs the WebGL widget with a ggplot2 reference panel so a package user can
judge how much of the analytical structure already survives the current
translation path.
