---
title: "Surface and mesh rendering with ggWebGL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Surface and mesh rendering with ggWebGL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

if (file.exists("DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(".", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else if (file.exists("../DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else {
  library(ggWebGL)
}

example_candidates <- c(
  "inst/examples/htmlwidget/surface-gallery.R",
  file.path("..", "inst", "examples", "htmlwidget", "surface-gallery.R"),
  system.file("examples", "htmlwidget", "surface-gallery.R", package = "ggWebGL")
)
example_candidates <- example_candidates[nzchar(example_candidates) & file.exists(example_candidates)]
if (!length(example_candidates)) {
  stop("Could not find surface-gallery.R")
}
sys.source(example_candidates[[1L]], envir = knitr::knit_global())
```

# Purpose

`ggWebGL` separates structured surfaces from unstructured meshes. Structured
surfaces are rectilinear grids that can carry height colormaps, normals,
wireframes, and optional contour overlays. Meshes are indexed triangle geometry
with scalar coloring, wireframe edges, and renderer-owned face or vertex ids.

Both examples below use small deterministic in-memory data. They are intended to
show the renderer contracts, not to report performance numbers.

# Structured Surface

The `volcano` matrix is converted with `surface_matrix()` and rendered as a
first-class surface primitive. The widget uses a perspective orbit view and the
`surface_height_colormap` shader.

```{r volcano-surface, out.width='100%'}
surface_gallery_volcano_widget(height = 460)
```

# Unstructured Mesh

The mesh example uses explicit vertices and triangle indices. Scalar values are
stored on vertices and displayed with `mesh_scalar_colormap`; face ids remain
available to the renderer for hover or selection payloads.

```{r scalar-mesh, out.width='100%'}
surface_gallery_mesh_widget(height = 460)
```

# Notes

The core package keeps rendering in the browser through WebGL. Any expensive
terrain preprocessing, mesh simplification, or device-specific acceleration is
left to upstream analysis code or optional companion tooling.
