---
title: "Temporal and 3D trajectories with ggWebGL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Temporal and 3D trajectories with ggWebGL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

if (file.exists("DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(".", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else if (file.exists("../DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else {
  library(ggWebGL)
}

example_candidates <- c(
  "inst/examples/htmlwidget/temporal-trajectories.R",
  file.path("..", "inst", "examples", "htmlwidget", "temporal-trajectories.R"),
  system.file("examples", "htmlwidget", "temporal-trajectories.R", package = "ggWebGL")
)
example_candidates <- example_candidates[nzchar(example_candidates) & file.exists(example_candidates)]
if (!length(example_candidates)) {
  stop("Could not find temporal-trajectories.R")
}
sys.source(example_candidates[[1L]], envir = knitr::knit_global())
```

# Purpose

`ggWebGL` can render ordered trajectory/path layers with timeline metadata. The
examples below use deterministic in-memory data and browser-side WebGL through
`htmlwidgets`; they do not download data or report performance numbers.
The widgets autoplay by default and can be paused, scrubbed, or reset with the
timeline controls. If a browser throttles a background tab, pressing **Play**
resumes the timeline.

**Status.** Trajectory paths, trajectory shader encodings, exact/cumulative
timeline controls, and Shiny timeline messages are `Experimental` public APIs.
They are suitable for renderer evaluation and small examples, but should not be
described as stable performance or animation guarantees.

The relevant renderer pieces are:

- `geom_path3d_webgl()` for ordered paths with optional `z`, `frame`, and `time`
  aesthetics.
- `animation_spec()` / `ggwebgl_timeline()` for exact or cumulative timeline
  metadata.
- `trajectory_age`, `trajectory_velocity`, and `trajectory_direction` shader
  modes for renderer-side trajectory visual encodings.
- `input$<outputId>_timeline` and `updateGgWebGLTimeline()` for Shiny timeline
  integration.

# 2D Cumulative Trajectory

This spiral is rendered as an ordered path with `frame` metadata and cumulative
timeline filtering. The `trajectory_age` shader makes the time prefix visible
as the timeline is scrubbed.

```{r spiral-trajectory, out.width='100%'}
temporal_spiral_widget(height = 460)
```

# 3D Trajectory

The helix uses the same path contract, but supplies `z` and `time` values. The
view is opt-in 3D with an orbit camera.

```{r helix-trajectory, out.width='100%'}
temporal_helix_widget(height = 460)
```

# Velocity and Direction Encodings

Trajectory velocity and direction are renderer-side visual encodings computed
from the serialized path coordinates and timeline values. They do not add
application semantics to the data.

```{r velocity-trajectory, out.width='100%'}
temporal_velocity_widget(height = 460)
```

```{r direction-trajectory, out.width='100%'}
temporal_direction_widget(height = 460)
```

# Exact Frame Samples

Exact filtering is useful for point-like animated samples. This example shows
only the samples whose `frame` equals the selected timeline value.

```{r exact-particles, out.width='100%'}
temporal_exact_particles_widget(height = 460)
```

# Shiny Timeline Integration

In Shiny, the browser emits the current timeline state as
`input$<outputId>_timeline`. Server code can update the browser-side timeline
with `updateGgWebGLTimeline()`.

```{r shiny-timeline-snippet, eval = FALSE}
library(shiny)
library(ggWebGL)

ui <- fluidPage(
  actionButton("restart", "Reset"),
  ggWebGLOutput("plot", height = "420px"),
  verbatimTextOutput("timeline_state")
)

server <- function(input, output, session) {
  output$plot <- renderGgWebGL({
    temporal_helix_widget(height = 460)
  })

  observeEvent(input$restart, {
    updateGgWebGLTimeline(session, "plot", index = 1L, playing = FALSE)
  })

  output$timeline_state <- renderPrint({
    input$plot_timeline
  })
}

if (interactive()) {
  shinyApp(ui, server)
}
```

The repository also includes a compact optional app at
`inst/examples/shiny/timeline-control-demo.R`.

# Limitations

Missing values and group boundaries break paths. Cumulative timeline filtering
draws complete visible prefixes; exact path filtering does not interpolate
partial hidden segments. Velocity and direction are renderer-side visual
encodings. Shiny timeline events report renderer time/frame state; downstream
applications remain responsible for interpreting those values.
